### 2020-12-10 Fig3_recreation.Rmd
recreating fig 3 progress

---
title: "fig3"
author: "Jonathan Olavarria"
date: "12/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
co_sim1 <- loadRData('co_redist_mcmc1.RData')
co_sim2 <- loadRData('co_redist_mcmc2.RData')
co_sim3 <- loadRData('co_redist_mcmc3.RData')
# TODO: are these the correct columns to use from Colorado data?
co_sim1_seg <- redist.segcalc(co_sim1, co$USH18R, co$TOTPOP)
co_sim2_seg <- redist.segcalc(co_sim2, co$USH18R, co$TOTPOP)
co_sim3_seg <- redist.segcalc(co_sim3, co$USH18R, co$TOTPOP)

rep.seg.mcmc <- data.frame(co_sim1_seg)
names(rep.seg.mcmc) <- "dissim"
nopop <- vector(mode = "list")
nopop$sim.hard <- rep.seg.mcmc
```

```{r, include = TRUE}
xlim <- c(0.1, max(nopop$sim.hard$dissim) + .01)
ylim <- c(0, max(density(nopop$sim.hard$dissim)$y))

## No constraint
plot(density(nopop$sim.hard$dissim),
     xlab = "Republican Dissimilarity Index",
     main = "",
     xlim = xlim,
     ylim = ylim,
     cex.lab = 1.6,
     cex.axis = 1.7)
polygon(density(nopop$sim.hard$dissim), col = "grey", border = "grey")

```