### 2020-12-13 paper.Rmd
Results section update

---
title: "MCMC Redistricting"
author: "Dawson Eliasen, Jonathan Olavarria, Henrique Rio"
date: "12/11/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

# Motivation

## Introduction

  Fair elections are a integral part of democracy for without it, the will of the people might not be respected and if that happens we no longer have a democracy. However, despite the huge importance of the elections in the US, there are still ways that a party can negatively impact the election by gaining an unfair advantage. One of the ways that this can happen, is gerrymandering where in order to gain a political advantage in a state by manipulating district boundaries. This issues has come before the supreme court, however it was ruled that it was up to the congress and the states to deal with it.

  Due to the scale of this issue, the paper by Fifield et al, proposes the use of a Markov Chain Monte Carlo method to automate the redistricting process, despite having a few Monte Carlo algorithms existing none of them have strong theoretical background. Fifield et al algorithm proposes to solve the automation redistricting by formulating it as a Graph Cut problem, where the MCMC samples the districts plans from arbitrary distributions over the set of n contiguous districts, they also implemented equal population constraint, which is a main requirement in the real-world problem.

## What is Redistricting?
 "In the United States is the process of drawing electoral district boundaries. A congressional act passed in 1967 requires that representatives be elected from single-member districts, except when a state has a single representative, in which case one state-wide at-large election be held." (Wikipedia). Redistricting also has to follow 6 different criteria, which are as follows.

Redistricting criteria:
    - compactness
    - contiguity
    - equal population
    - preservation of existing political communities
    - partisan fairness
    - racial fairness
  The compact criteria pertains to the size of each district where one district cannot be of a enormous size and another one be really small, they all have to be more or less similar in size. The contiguity constraint, stipulates that all that all of the district have to be continuous in area in the sense that one district cannot cut through the middle of the other one. The equal population constraint, e

## Gerrymandering
- What is Gerrymandering?
    - Is the act of actively redrawing district lines so that it tilts as many districts as possible in favor of a party, in order for a party to cement themselves in power.
    - e.g. A controversial redistricting that happened in Texas in 2003, where          republicans went from 15 seats in 2002 to 21 in 2004 (democrats got 11).
    - After this the case was brought to the supreme court, where it was ruled that the supreme court could not address gerrymandering, which leaves to congress and states to deal with it.

## Automated Redistricting
- Researchers have been trying to create a redistricting algorithm since the 1970s
- It began as a optimization problem, where the goal is trying to find the optimal redistricting plan.
- Due to the geographical dependence of redistricting in a state, it is essential to determine the distribution of redistricting plans under the redistricting criteria.
- Simulation Methods are more appropriate since it allows scholars to answer these questions,by  simulating different plans outcomes under different constraints.
- Yet only a few exist so far, using the same Monte Carlo simulation algorithm.
- We chose to reproduce a paper by Fifield from Harvard et. al., *Automated Redistricting Simulation Using Markov Chain Monte Carlo*

# Methodology

## Redistricting Problem Representation
Redistricting is represented as a “graph cut” problem, where a state is represented as a graph $G$, and each voting precinct $i$ is a node in the set of nodes $V$. There is an undirected edge $i \sim j$ connecting nodes $i, j$ if the precincts $i, j$ are contiguous (share a border). Then, we get voter districts by partitioning the set of nodes $V$. This leaves subgraphs that represent the congressional districts. Since districts are formed by “cutting” edges, we can think of redistricting as a “graph cut” problem. This is how the MCMC methods generate new samples of redistricting plans.

## Metropolis-Hastings
The Metropolis-Hastings algorithm is a Monte Carlo Markov Chain algorithm based on the simple accept-reject algorithm, used to simulate redistricting plans.

- Define a function $f(x)$, proportional to the pdf of the desired distribution
- Define a proposal distribution $g(x | y)$ that provides a new sample $x$ given the previous sample $y$
- Start with a valid sample, $x_0$
- At each iteration:
    - Generate a “candidate” sample $x^*$ from $g$ (perturb $x_i$)
    - Accept $x^*$ with probability equal to $\frac{f(x^*)}{f(x_i)}$

## Metropolis-Hastings Iteration Example
To simulate redistricting, “perturbation” looks like swapping precincts along existing district borders. \ref{fig:MCMC example} illustrates the perturbation process for an example sample in a Metropolis-Hastings chain. Perturbation works by identifying nodes that exist on congressional district borders and proposing swaps along the borders. This produces a perturbation of the previous sample which is accepted with acceptance ratio

$$
\alpha(\pi', \mathbf{CP} | \pi) = \min \Big(1,
    \Big(\frac{|B(\mathbf{CP}, \pi)|}{|B(\mathbf{CP}, \pi'|}\Big)^R 
    \frac{F(|B(\mathbf{CP}, \pi)|)(1- q)^{C(\pi', \mathbf{V_{CP}})}}{F(|B(\mathbf{CP}, \pi')|) (1- q)^{C(\pi, \mathbf{V_{CP}})}} \cdot
    \frac{g(\pi')}{g(\pi)} 
\Big)
$$

where $\pi$ is the current sample, $\pi'$ is the perturbed sample, $\mathbf{CP}$ is the set of connected components formed by $\pi$, and $C$ is a graph-cut function.

## `redist` Package
Fifield et. al. use a package called `redist` which implements this Metropolis-Hastings algorithm for simulating redistricting. `redist.mcmc` is a function for generating redistricting samples, and requires precinct adjacency matrix, population data, and assignments to a valid set of congressional districts as function inputs. The authors of the paper shared some code for reproduction, but the code was limited to generating figures and it was not generalizable to different data, so we re-implemented the simulation code to generate redistricting simulations for the state of Colorado.

## Data
Fifield et. al. use Pennsylvania shapefiles and election data from the Harvard Election Archive. However, the Harvard Election Archive Colorado dataset did not include congressional district assignments. Instead, we used a shapefile assembled by MGGG which was constructed from election data from Colorado Secretary of State’s Office, 2010 Census data from the U.S. Census Bureau Data API, and precinct borders from U.S. Census Bureau TIGER/Line shapefile repository.

## Simulation
Using the `sf` package, we read the Colorado shapefiles and calculated a precinct adjacency matrix with a package called `spdep`. Then, we used `redist.mcmc` to generate 3 MCMC chains of 10,000 samples each - this took a few hours. We initially tried 3 MCMC chains of 100,000 samples (this is what the authors of the paper do) - but this took to long and required more computing power than was available. We also tried applying a “compactness” constraint, but it requires providing a distance matrix of the nodes in the graph, which took too long to compute. Finally, we evaluate bias in the simulations using `redist.segcalc` and a modified version of the code shared by the authors of the paper.

# Results

## Starting Map (current districts)
```{r, include=FALSE}
library(tidyverse)
library(plyr)
library(sf)
library(RColorBrewer)
source('plotting_functions.R')
```

```{r, echo=FALSE, fig.cap="Current congressional districts in Colorado.\\label{fig:map}"}
co <- read_sf(dsn='co_precincts', layer='co_precincts')
co <- co %>% arrange(CD116FP)
from <- unique(co$CD116FP)
to <- seq_len(length(unique(co$CD116FP)))
co$CD116FP <- co$CD116FP %>% mapvalues(from, to)

ggplot(co, aes(fill=CD116FP)) +
    geom_sf(size=0) +
    scale_fill_brewer(palette='Set2')
```

In figure 1 above we see the current district map for the state of colorado. The districts are distinguished by the different colors and the legend on the right. The opaque white lines show the precinct lines (there are over 3000 precincts in the state of Colorado), as mentioned in the intorduction gerrymandering is the process of moving precincts from one district to another to sway political bias in certain districts or in an entire state. Each district shown above satisfies the equal population constraint stating that every district must have approximately the same population. At the moment politicians are doing this process by hand looking at each precinct's voting history and moving them around to benefit their political party. Redistricting happens once every 10 years after every census. The current district layout exhibits democratic bias which is as expected because democrats controlled the state legislature in 2010 which was a census year.

## MCMC Summary

```{r, include=FALSE}
loadRData <- function(f) {
    load(f)
    get(ls()[ls() != 'f'])
}
```

```{r, include=FALSE}
co_sim1 <- loadRData('co_redist_mcmc1.RData')
co_sim2 <- loadRData('co_redist_mcmc2.RData')
co_sim3 <- loadRData('co_redist_mcmc3.RData')
co_sim1_seg <- redist.segcalc(co_sim1, co$USH18R, co$TOTPOP)
co_sim2_seg <- redist.segcalc(co_sim2, co$USH18R, co$TOTPOP)
co_sim3_seg <- redist.segcalc(co_sim3, co$USH18R, co$TOTPOP)
co_sim <- cbind(co_sim1$partitions[,!is.na(co_sim1_seg)],
                co_sim2$partitions[,!is.na(co_sim2_seg)], 
                co_sim3$partitions[,!is.na(co_sim3_seg)])
xax <- 1 - c(co_sim1$distance_original[!is.na(co_sim1_seg)],
             co_sim2$distance_original[!is.na(co_sim2_seg)],
             co_sim3$distance_original[!is.na(co_sim3_seg)])
co_sim_bias <- calc_sim_bias(co_sim, co$GOV18D, co$GOV18R)
```

```{r, echo=FALSE, fig.cap="Redistricting samples, percent of precincts changed vs. bias.\\label{fig:mcmc-summary}"}
plot_sim_bias(xax, co_sim_bias)
```

## Minimum Bias Map
```{r, echo=FALSE, fig.cap="Redistricting sample with minimum bias.\\label{fig:min-bias-map}"}
plot_min_bias(co_sim, co_sim_bias, co, xax)
```